---
title: "HW6"
output: github_document
author: Serena (sjt2164)
---
```{r}
library(tidyverse)
```

##Problem 1

```{r weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r summary, include=FALSE}
weather_df
#365 observations, 6 variables
#chr: name, id = Central Park data only
#date: date = 2017 data from Jan 1st to Dec 31st.
#dbl: prcp, tmax, tmin

summary(weather_df)
#no NAs for pcrp, tmax, tmin
```

_Prompt_

* `tmax` as the response 
* `tmin` as the predictor
* interested in distribution of two quantities estimated from:
  * ğ‘ŸÌ‚ 2
  * log(ğ›½Ì‚ 0âˆ—ğ›½Ì‚ 1)
  
* Use 5000 bootstrap samples
* For each bootstrap sample, produce estimates of these two quantities. 
* Plot the distribution of your estimates, and describe these in words. 

* Using the 5000 bootstrap estimates, identify the 2.5% and 97.5% quantiles to provide a 95% confidence interval for ğ‘ŸÌ‚ 2 and log(ğ›½Ì‚ 0âˆ—ğ›½Ì‚ 1)
* Use broom::glance() for extracting ğ‘ŸÌ‚ 2 from a fitted regression
* Use broom::tidy() (with some additional wrangling) to compute log(ğ›½Ì‚ 0âˆ—ğ›½Ì‚ 1)

