---
title: "HW6"
output: github_document
author: Serena (sjt2164)
---

This homework uses the `tidyverse` library with `set.set(1)` for reproducibility.

```{r, include = FALSE}
library(tidyverse)

set.seed(1)
```

## Problem 1

```{r weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

#Central Park data for year 2017 with info on precipitation, max and min temp
```

```{r summary, include=FALSE}
weather_df
#365 observations, 6 variables
#chr: name, id = Central Park data only
#date: date = 2017 data from Jan 1st to Dec 31st.
#dbl: prcp, tmax, tmin

summary(weather_df)
#no NAs for pcrp, tmax, tmin
```

* For each boostrap sample, interested in distribution of :
  * 𝑟̂ 2
  * log(𝛽̂ 0∗𝛽̂ 1)

```{r bootstrap}
Bootstrap = weather_df |> 
  modelr::bootstrap(n = 5000) |> #5000 bootstrap samples 
  mutate(
    models = map(strap, \(df) lm(tmax ~ tmin, data = df) ), # y = tmax, x = tmin
    results = map(models, broom::tidy), #for log betas
    r_squared = map(models, broom::glance)) |> #shows R^2
   # log_betas = log(intercept * 
  unnest(results, r_squared) |> 
  janitor::clean_names() |>
  select(id, term, estimate, r_squared) |>
   mutate(
    term = ifelse(term == "(Intercept)", "intercept", term)) |>
  pivot_wider(
    names_from = "term", 
    values_from = "estimate") |>
  mutate(
    log_beta0xbeta1 = log(intercept * tmin)) |> #log of the product of beta0 and beta1
  select(id, r_squared, log_beta0xbeta1)

Bootstrap
```

##### Distribution of 𝑟̂ 2 estimates

```{r estimate plots}
Bootstrap |>
  ggplot(aes(x = id, y = r_squared)) + 
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm") +
  labs(title = "R squared distribution for 5000 bootstrap samples",
       x = "Boostrap Sample Id",
       y = "R squared") 
  
```

##### Distribution of log(𝛽̂ 0∗𝛽̂ 1) estimates

```{r}
Bootstrap |>
  ggplot(aes(x = id, y = log_beta0xbeta1)) + 
  geom_point(alpha = 0.5) +
  stat_smooth(method = "lm") +
  labs(title = "Log(beta0 * beta1) distribution for 5000 bootstrap samples",
       x = "Boostrap Sample Id",
       y = "Log(beta0 * beta1") 

```

Both plots indicate an uniformly distribution of the estimates. The distribution of the R squared estimates has a higher density around 0.90 and 0.92. In comparison, the  log(𝛽̂ 0∗𝛽̂ distribution is more concentrated between 2.00 and 2.05. 
       
       
##### Confidence Intervals

```{r}
CI_estimates = Bootstrap |> 
  summarize(
    R2_ci_lower = quantile(r_squared, 0.025), 
    R2_ci_upper = quantile(r_squared, 0.975),
    logBetas_ci_lower = quantile(log_beta0xbeta1, 0.025), 
    logBetas_ci_upper = quantile(log_beta0xbeta1, 0.975)) 

CI_estimates
```


Using the 5000 bootstrap estimates, we find that the 95% confidence interval for 𝑟̂ 2 is (0.894, 0.927).

The 95% confidence interval for log(𝛽̂ 0∗𝛽̂ 1) is (1.964, 2.059).


## Problem 2: Homicides in 50 large U.S. cities 

```{r Homicide data}
#Import from github repo, read article.
```

* Create a `city_state` variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved (ex. `solved`: y/n)

* Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO (they don’t report victim race) 
* Omit Tulsa, AL (is a data entry mistake)
* Limit your analysis to those for whom `victim_race` is `white` or `black`. 
* Be sure that `victim_age` is numeric.

_*_Baltimore, MD_

* For the city of Baltimore, MD, use the `glm` function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. 
* Save output of `glm` as an R object; apply the `broom::tidy` to this object
* Obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

_Glm for the Cities_

* Now run glm for each of the cities in your dataset
* Extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. 
* Do this within a “tidy” pipeline, making use of `purrr::map`, list columns, and `unnest` as necessary to create a dataframe with estimated ORs and CIs for each city.

*Plot*

```{r plot 1}

```

* Shows the estimated ORs and CIs for each city on plot
* Organize cities according to estimated OR
* Comment on the plot

## Problem 3: Effects of variables on a child’s birthweight

```{r bwt}

bwt_df = read_csv("./data/birthweight.csv", na = c("NA", ".", ""))

#4342 observations, 20 variables

```

Dataset has ~4000 children and includes the following variables:

* `babysex`: baby’s sex (male = 1, female = 2)
* `bhead`: baby’s head circumference at birth (centimeters)
* `blength`: baby’s length at birth (centimeteres)
* `bwt`: baby’s birth weight (grams)
* `delwt`: mother’s weight at delivery (pounds)
* `fincome`: family monthly income (in hundreds, rounded)
* `frace`: father’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other, 9 = Unknown)
* `gaweeks`: gestational age in weeks
* `malform`: presence of malformations that could affect weight (0 = absent, 1 = present)
* `menarche`: mother’s age at menarche (years)
* `mheigth`: mother’s height (inches)
* `momage`: mother’s age at delivery (years)
* `mrace`: mother’s race (1 = White, 2 = Black, 3 = Asian, 4 = Puerto Rican, 8 = Other)
* `parity`: number of live births prior to this pregnancy
* `pnumlbw`: previous number of low birth weight babies
* `pnumgsa`: number of prior small for gestational age babies
* `ppbmi`: mother’s pre-pregnancy BMI
* `ppwt`: mother’s pre-pregnancy weight (pounds)
* `smoken`: average number of cigarettes smoked per day during pregnancy
* wtgain`: mother’s weight gain during pregnancy (pounds)

#### Data Cleaning
```{r bwt clean}
#clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```

#### Regression Model

```{r Model}

```

Propose a regression model for birthweight. 
* Model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. 
* Describe modeling process

```{r plot 2}

```

* Show a plot of model residuals against fitted values – use `add_predictions` and `add_residuals` in making this plot.

#### Compare your model to two others:

* One using length at birth and gestational age as predictors (main effects only)

* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

Make this comparison in terms of the cross-validated prediction error; use `crossv_mc` and functions in `purrr` as appropriate.

Note that although we expect your model to be reasonable, model building itself is not a main idea of the course and we don’t necessarily expect your model to be “optimal”.

